Towards something :-D

### 2015-03-25 paper/nodalida2015.tex
more text

\title{Towards ....}

@@

\section{Introduction}
Constraint Grammar (CG) was first introduced by \cite{karlsson1995constraint}. 
It is a tool for disambiguating output by morphological analyser.

Our implementation is inspired by approaches like \cite{lager98,lager_nivre01},
but we try to emulate the CG formalism more closely, and use the same rule files.

@@

Application of logic to POS tagging or shallow parsing isn't new;
\cite{lager98,lager_nivre01} present constraint rules as 
a disjunctive logic program.

\cite{lindberg_eineborg98ilp,asfrent14}

% What we do better: implement something that takes existing CG rules and applies them to text, trying to mimic/outperform current CG implementations.

\cite{lager01transformation} showed rules where ordering didn't matter so much in the end.

@@

SAT solving is based on a technique called unit propagation:
a set of complex Boolean clauses is simplified, starting from unit
clauses, which consist of just a single variable, and working up to a
solution, where all variables in the set have a True or False value.

SAT solving is used for many applications where constraints must be
satisfied, such as software and hardware verification.

Let us demonstrate the approach with the example sentence 
\emph{the bear sleeps}. After applying morphological tagger, we get the following analyses:

\begin{verbatim}
"<the>"
        "the" det
"<bear>"
        "bear" n sg
        "bear" v pl
"<sleeps>"
        "sleep" n pl
        "sleep" v sg p3
\end{verbatim}

We make each analysis a variable; let us call them \texttt{d1}, \texttt{n2}, \texttt{v2}, \texttt{n3} and \texttt{v3} by the proposed POS tag and the position in the sentence.
For instance, \texttt{n2} is a hypothesis that the correct analysis 
for the word form \emph{bear}' is \texttt{"bear" n pl}.
Before applying rules, we must make sure that in each position
at least one of the analyses is true. We add our first disjunctions:

\begin{itemize}
\item [] $d1$ \\ $n2 \vee v2$ \\ $n3 \vee v3$
\end{itemize}

Let us add two rules:
\begin{itemize}
\item [] \texttt{REMOVE v IF -1 det} \\ 
            \texttt{REMOVE n IF -1 n} 
%             $d1 \Rightarrow \neg v2$ 
%\item [] \texttt{REMOVE n IF -1 n} \\
%              $n2   \Rightarrow \neg n3$

\end{itemize}

The rules are applied as implications: \emph{if ``the'' is a determiner, then ``bear'' cannot be a verb}. 
The antecedent is

@@

The condition of being unambiguously tagged becomes less necessary as well.
We add many clauses, knowing that some of them end up being false.

@@

Ordering and conflicts : the most important differences here

%\cite{koskenniemi92}
%\begin{quote}Rules in the CG formalism are typically [--] executed as successive groups.
%In finite-state syntax, rules are logically unordered.\end{quote}

@@

\subsection{Other experiments}

@@

\paragraph{Less rules needed to disambiguate.}
The following example illustrates the interaction of rules to produce the desired outcome. We have an ambiguous passage in Spanish, with the following analyses:

@@

We took small constraint grammars (<500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg. We compared the result of using VISLCG3 for the same task; table~\ref{table:results} shows the differences in results in both sentence and word level.

\begin{table}
  \centering
  \begin{tabular}{|r|r|r|r|r|}
     \hline
    \textbf{Text} & \textbf{Diff words} & \textbf{Diff sentences} \\
    \hline
	Pride and Prejudice & todo & 5457/7667 \\  
    \hline
       Don Quijote & todo & 3k/6k  \\ 
     \hline
  \end{tabular}
  \caption{Results for comparing SATCG and VISLCG3.}
  \label{table:results}
\end{table}

These tendencies show also in a larger context with more complex rules.

>>>>

\title{Towards something :-D}

@@

\section{Introduction}
Constraint Grammar (CG) was first introduced by \cite{karlsson1995constraint}. 
It disambiguates output by morphological analyser, 
and has been used for many tasks in computational linguistics, such as POS tagging,
surface syntax and machine translation.
The current state-of-the-art parser VISL CG-3 is implemented (...);
in addition there are a number of finite-state based implementations \cite{nemeskey14}.

We represent CG as a Boolean satisfiability problem using the 
open-source SAT solver MiniSat \cite{een04sat}.
Our implementation is inspired by approaches like \cite{lager98,lager_nivre01},
but we try to emulate the CG formalism more closely than the previous logic-based works.

% NP-complete -- overkill for CG? But gives other benefits.

@@

Application of logic to POS tagging or shallow parsing isn't new;
\cite{lager98,lager_nivre01} present constraint rules as 
a disjunctive logic program.
\cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
\cite{lager01transformation} discusses ordering of the CG rules.

Works from the late 90s and early 00s mention that the method is very slow;
however, advances in SAT solving technique

@@

SAT solving is based on a technique called unit propagation:
a set of complex Boolean clauses is simplified, starting from unit
clauses, which consist of just a single variable, and working up to a
solution, where all variables in the set have a True or False value.
We model an ambiguous sentence as a set of clauses, where each analysis is a variable, 
and constraint rules form clauses operating on those variables.

\begin{itemize}
\item The condition of being unambiguously tagged becomes less necessary.
We add many clauses, knowing that some of them end up being false.
\item Rules are unordered. 
\item Rules interact with each other and disambiguate more than same rules in the standard implementation.
\end{itemize}

% Let us demonstrate the approach with the example sentence 
% \emph{the bear sleeps}. After applying morphological tagger, we get the following analyses:

% \begin{verbatim}
% "<the>"
%         "the" det
% "<bear>"
%         "bear" n sg
%         "bear" v pl
% "<sleeps>"
%         "sleep" n pl
%         "sleep" v sg p3
% \end{verbatim}

%We make each analysis a variable; let us call them \texttt{d1}, \texttt{n2}, \texttt{v2}, \texttt{n3} and \texttt{v3} by the proposed POS tag and the position in the sentence.
%For instance, \texttt{n2} is a hypothesis that the correct analysis 
%for the word form \emph{bear}' is \texttt{"bear" n pl}.
%Before applying rules, we must make sure that in each position
%at least one of the analyses is true. We add our first disjunctions:

%\begin{itemize}
%\item [] $d1$ \\ $n2 \vee v2$ \\ $n3 \vee v3$
%\end{itemize}

%Let us add two rules:
%\begin{itemize}
%\item [] \texttt{REMOVE v IF -1 det} \\ 
%            \texttt{REMOVE n IF -1 n} 
%\end{itemize}

%The rules are applied as implications: \emph{if ``the'' is a determiner, then ``bear'' cannot be a verb}.

@@

@@

Unordered rules are not new: finite-state implementations of CG \cite{koskenniemi92} also have logically unordered rules.

@@

@@

\subsection{Rule interaction}

The following example illustrates the interaction of rules to produce the desired outcome. We have an ambiguous passage in Spanish, with the following analyses:

@@

We took small constraint grammars (\textless{}500 rules) for Spanish and English, and disambiguated a number of texts from Project Gutenberg. We compared the result of using VISLCG3 for the same task; table~\ref{table:results} shows the differences in results in both sentence and word level.

\begin{table}
  \centering
  \begin{tabular}{|r|r|r|r|r|}
     \hline
    \textbf{Text} & \textbf{Diff words} & \textbf{Diff sentences} \\
    \hline
	Pride and Prejudice & todo & 5k/7k \\  
    \hline
       Don Quijote & todo & 3k/6k  \\ 
     \hline
  \end{tabular}
  \caption{Results for comparing CGSAT and VISLCG3.}
  \label{table:results}
\end{table}

The tendency of needing less rules to disambiguate shows also in a larger context with more complex rules.
In many of the differences, CGSAT disambiguated more. (TODO counts)
